---
title: "MA677 FINAL PROJECT"
author: "Kecheng Liang"
date: "May 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse,knitr,pwr,fitdistrplus)
```

#Statistics and the Law
two sample t-test
Null hypothesis: refuse rate of minority applicant is the same as that of white applicant  
Alternative hypothesis: refuse rate of minority applicant is higher that of white applicant
```{r}
acorn<-read.csv("acorn.csv")
test1 <- var.test(acorn$MIN,acorn$WHITE)
test1
test2 <- t.test(acorn$MIN, acorn$WHITE, alternative = "greater",var.equal = FALSE)
test2
```
First, I did the F test to compare two variances. Since p-value = 0.02993, reject the null hypothesis, the two variances are not same. Then I did the two sample t-test. The result shows that the p-value = 2.979e-07, reject the null hypothesis. The refuse rate of minority applicant is higher that of white applicant.

#Comparing Suppliers
Chi-square test
Null hypothesis: all three schools produces the same quality  
Alternative hypothesis: at least one of these three schools produces different quality
```{r}
df <- matrix(c(12,8,21,23,12,30,89,62,119),nrow = 3,byrow = FALSE)
colnames(df) <- c("Dead Bird","Display Art","Flying Art")
rownames(df) <- c("Area 51","BDV","Giffen")
chisq.test(df)
```
By doing the chi-square test, p-value = 0.8613 which is greater than 0.05. Therefore, we fail to reject the null hypothesis that all three schools produces the same quality.

#How deadly are sharks?
```{r}
shark<-read.csv("sharkattack.csv")
sharkdf <- shark %>% 
            filter(Country == "United States" | Country == "Australia") %>%
            filter(Type == "Provoked" | Type == "Unprovoked")
table <- table(droplevels(sharkdf)$Country,droplevels(sharkdf)$Type)
kable(prop.table(table,margin = 1))
chisq.test(table)
sharkdf2 <- shark %>% 
            filter(Country == "United States" | Country == "Australia") %>%
            filter(Fatal == "Y" | Fatal == "N")
table2 <- table(droplevels(sharkdf2)$Country,droplevels(sharkdf2)$Fatal)
kable(prop.table(table2,margin = 1))
chisq.test(table2)
pwr.chisq.test(w = ES.w2(prop.table(table2)), N=879+318+1795+217, df = 1, sig.level = 0.05)
```
I did 2 chi-square tests. The first one is try to compare the provoked and unprovoked between US and Australia. The result shows that p-value = 0.07907, and we do not reject the null hypothesis. Therefore, there is no association betweeen two variables. The p-value from second one is < 2.2e-16. Therefore, we reject the null hypothesis, there is association between fatal and country. The attack in Australia is much more deadly. The power is 1.

#Power analysis
In the book, it said that the hypothetical parameters of this binomial distribution doesn not provide a scale of equal units of detectability. Arcsin transformation could solve the problem that falling into one side of the range.symbol = 2 arcsin root(P).

#Estimators


#Rain in Southern Illinois
```{r}
ill60 <- read.table("ill-60.txt")
year60<-as.numeric(as.array(ill60[,1]))
ill61 <- read.table("ill-61.txt")
year61<-as.numeric(as.array(ill61[,1]))
ill62 <- read.table("ill-62.txt")
year62<-as.numeric(as.array(ill62[,1]))
ill63 <- read.table("ill-63.txt")
year63<-as.numeric(as.array(ill63[,1]))
ill64 <- read.table("ill-64.txt")
year64<-as.numeric(as.array(ill64[,1]))
plotdist(year60)
plotdist(year61)
plotdist(year62)
plotdist(year63)
plotdist(year64)
#sum of year 60
sum(year60)
#sum of year 61
sum(year61)
#sum of year 62
sum(year62)
#sum of year 63
sum(year63)
#sum of year 64
sum(year64)
```
Year 1961 is wetter based on the calculation for total amount of rainfall. However I cannot find any obvious different in those five distributions. Most of the rainfall are concentrated at the left side of data.

```{r}
years <- c(year60,year61,year62,year63,year64)
gammadist <- fitdist(years, "gamma")
plot(gammadist)
summary(gammadist)
```
From the output, we can see that it fits well. Changnon and Huff are right about using gamma distribution.

```{r}
gamma2 <- fitdist(years, "gamma",method = "mme")
gamma2d <- bootdist(gamma2)
summary(gamma2d)

gamma3 <- fitdist(years, "gamma",method = "mle")
gamma3d <- bootdist(gamma3)
summary(gamma3d)
```
Compare those two methods, mle has narrower CI. Therefore, I would choose mle as the estimator because it has the lower variance.

#Decision theory
